---
title: "ex_6_soln"
output: html_document
date: "2026-06-11"
author: "Suvansh Shukla"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r cars}
library(tidyverse)
library(flexclust)
library(biclust)
library(reshape)
library(ggplot2) 
library(GGally)
library(hexbin)

# get girth, height & volume from default tree dataset
h <- trees$Height
g <- trees$Girth
v <- trees$Volume

plot(g, h, main="Girth vs Height", xlab="Girth", ylab="Height")
plot(g, v, main="Girth vs Volume", xlab="Girth", ylab="Volume")
plot(h, v, main="Height vs Volume", xlab="Height", ylab="Volume")
```

From the scatter plots, it seems pretty clear that "Girth" and "Volume" have some correlation.

```{r}
ggplot(trees, aes(x = Girth, y = Volume)) +
  geom_point(color = "blue", size = 2) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Scatter Plot with Pearson Correlation",
       x = "Girth", y = "Volume") +
  theme_minimal()

input_trees = trees[, c("Girth", "Height", "Volume")]
pairs(input_trees, labels = colnames(input_trees), pt.bg = c("red", "green3", "blue", "yellow2"),
      pch = 21, bty = "n", ncol = 2)
```

```{r}
data(trees)

#pairs(trees, upper.panel= main="Scatter plot matrix for trees dataset" ,pch=20, cex=1.5, col="steelblue")
Scatter_matrix <- ggpairs(trees, title="Scatter plot matrix for trees dataset", axisLabels="show")
Scatter_matrix

correlation_nums <- function(x, y, ...){
  r <- cor(x, y, method="pearson")
  txt <- format(c(r, 0.123456789), digits=2)[1]
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  text(0.5, 0.5, txt, cex=2)
}
pairs(trees, lower.panel=points, upper.panel=correlation_nums, main="Scatter plot matrix for Trees dataset")
```

## Question 2

This question involves using a biclustering algorithm and heatmaps.

```{r}
# Create bipartite graph data
game_matrix <- matrix(c(1, 0, 0, 1, 1, 1, 0,
                        0, 1, 1, 0, 1, 0, 0,
                        0, 0, 0, 0, 1, 0, 1,
                        1, 1, 0, 0, 0, 0, 0,
                        0, 0, 1, 0, 0, 0, 1,
                        0, 1, 0, 1, 1, 0, 0,
                        0, 0, 0, 0, 0, 1, 0),
                        nrow = 7,
                        ncol = 7,
                        byrow = TRUE)
colnames(game_matrix) <- c("The Witcher 3","Terraria","Hollow Knight","Cities: Skylines","GTA V","Divinity","XCOM 2")
rownames(game_matrix) <- c("Tarek", "Sam", "Ebele", "Kim", "Ali", "Aiko", "Parvati")

bc_res <- biclust(game_matrix, method=BCCC(), delta=0.5, number=5)
summary(bc_res)

image(1:ncol(game_matrix), 1:nrow(game_matrix), t(game_matrix), 
      col = c("white", "black"), axes = FALSE, xlab = "Games", ylab = "Students",
      main = "Heatmap 1: Original Game Data (Black = Plays)")
axis(1, at=1:ncol(game_matrix), labels=colnames(game_matrix), las=2, cex.axis=0.7)
axis(2, at=1:nrow(game_matrix), labels=rownames(game_matrix), las=1, cex.axis=0.7)

color_matrix <- matrix(0, nrow=nrow(game_matrix), ncol=ncol(game_matrix))

if(bc_res@Number > 0) {
  for(i in 1:bc_res@Number) {
    rows_in_bc <- bc_res@RowxNumber[, i]
    cols_in_bc <- bc_res@NumberxCol[i, ]
    
    color_matrix[rows_in_bc, cols_in_bc] <- i
  }
}

image(1:ncol(color_matrix), 1:nrow(color_matrix), t(color_matrix), 
      col = c("white", "tomato", "skyblue", "lightgreen"), axes = FALSE,
      xlab = "Games", ylab = "Students", main = "Heatmap 2: Bicluster Cell Grid")
axis(1, at=1:ncol(game_matrix), labels=colnames(game_matrix), las=2, cex.axis=0.7)
axis(2, at=1:nrow(game_matrix), labels=rownames(game_matrix), las=1, cex.axis=0.7)
grid(nx=7, ny=7, col="gray", lty="solid")
```

The core issue is overlapping biclusters.

## Question 3

```{r}
# Define data
# Set seed
set.seed(48)
get_rand <- function(d) {
  return(rnorm(1, mean = d, sd = 1))
}
m <- round(runif(50000, 0, 1)) * 2 + 1
point_data = tibble(x = sapply(m, get_rand), y = sapply(m, get_rand))

plot(point_data$x, point_data$y, main="X vs Y", xlab="X", ylab="Y")

x <- point_data$x
y <- point_data$y
hb <- hexbin(x, y, xbins=50)

plot(hb, main="HexBin plot")

```

The second "HexBin" plot is far better, it allows for "transparency", which makes the plot easier to read and understand. Changing the plot params helps in increasing/reducing the dark area in the hexbin plot.